---
title: "Preliminary Quality Control of Novartis DECAMP2 Broncial Brushings RNA-seq data"
date: "`r format(Sys.time(), '%d %B, %Y')`"

output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    fig_width: 9
    fig_height: 6
    theme: cosmo
geometry: margin=1in
---

## Preprocessing

```{r, message = FALSE, warning=FALSE}
# Load library:
library(plyr)
library(SummarizedExperiment)
library(readxl)
library(dplyr)
library(ggplot2)
library(pheatmap)
library("RColorBrewer")
library(DESeq2)
library(knitr)
library(kableExtra)

set.seed("12345")

# Load data:
se <- readRDS("/restricted/projectnb/pulmseq/Novartis/processing/work/28/da9dcb4ebb5f902c53fff1bc1ebdc2/Novartis_Gene_Expression.rds")
colnames(se) <- gsub(pattern = "_.*", "", colnames(se))

# Load annotation:

# Matches sample ID to kitnumber + Has batch info
annotManifest <- read_excel("/restricted/projectnb/pulmseq/Novartis/annotation/Novartis_RNASeq_run_manifest.xlsx")

# Matches sample ID to sample type/RIN (received from Jack, 8/24/20)
annotSampleEM0 <- read_excel("/restricted/projectnb/camplab/home/ykoga07/DECAMP/Novartis/Data/Annotation/RNA_data_for_Yusuke_08_24_20.xlsx", sheet = "EM0", skip = 2)

annotSampleEM5 <- read_excel("/restricted/projectnb/camplab/home/ykoga07/DECAMP/Novartis/Data/Annotation/RNA_data_for_Yusuke_08_24_20.xlsx", sheet = "EM5", skip = 2)

annotSampleEM6 <- read_excel("/restricted/projectnb/camplab/home/ykoga07/DECAMP/Novartis/Data/Annotation/RNA_data_for_Yusuke_08_24_20.xlsx", sheet = "EM6", skip = 2)

annotSampleEM8 <- read_excel("/restricted/projectnb/camplab/home/ykoga07/DECAMP/Novartis/Data/Annotation/RNA_data_for_Yusuke_08_24_20.xlsx", sheet = "EM8", skip = 3)

annotSampleEM15 <- read_excel("/restricted/projectnb/camplab/home/ykoga07/DECAMP/Novartis/Data/Annotation/RNA_data_for_Yusuke_08_24_20.xlsx", sheet = "EM15", skip = 3)

annotSampleEM16 <- read_excel("/restricted/projectnb/camplab/home/ykoga07/DECAMP/Novartis/Data/Annotation/RNA_data_for_Yusuke_08_24_20.xlsx", sheet = "EM16", skip = 3)

annotSampleN1 <- read_excel("/restricted/projectnb/camplab/home/ykoga07/DECAMP/Novartis/Data/Annotation/RNA_data_for_Yusuke_08_24_20.xlsx", sheet = "N1", skip = 3)

annotSampleN2 <- read_excel("/restricted/projectnb/camplab/home/ykoga07/DECAMP/Novartis/Data/Annotation/RNA_data_for_Yusuke_08_24_20.xlsx", sheet = "N2", skip = 3)

# Combine Jack's sample-based annotation file into one dataframe
annotSampleEM0 <- annotSampleEM0[, c(
  "RNA Sample ID", "DECAMP #",
  "Sample ID", "Kit Number", "Original Sample Type",
  "Isolation Date", "RIN", "DV200 (%)"
)]
colnames(annotSampleEM0) <- c(
  "RNA Sample ID", "DECAMP #",
  "Sample ID", "Kit Number", "Sample Type",
  "Isolation Date", "RIN", "DV200 (%)"
)

annotSampleEM5 <- annotSampleEM5[, c(
  "RNA Sample ID", "DECAMP #",
  "Patient ID", "Patient ID", "Sample Type",
  "Isolation Date", "RIN", "DV200 (%)"
)]
colnames(annotSampleEM5) <- c(
  "RNA Sample ID", "DECAMP #",
  "Sample ID", "Kit Number", "Sample Type",
  "Isolation Date", "RIN", "DV200 (%)"
)

annotSampleEM6 <- annotSampleEM6[, c(
  "RNA Sample ID", "DECAMP #",
  "Sample ID", "Kit Number", "Sample Type",
  "Isolation Date", "RIN", "DV200 (%)"
)]
colnames(annotSampleEM6) <- c(
  "RNA Sample ID", "DECAMP #",
  "Sample ID", "Kit Number", "Sample Type",
  "Isolation Date", "RIN", "DV200 (%)"
)

annotSampleEM8 <- annotSampleEM8[, c(
  "RNA Sample ID", "DECAMP #",
  "Sample ID", "Original Sample ID", "Sample Type",
  "Isolation Date", "RIN", "DV200 (%)"
)]
colnames(annotSampleEM8) <- c(
  "RNA Sample ID", "DECAMP #",
  "Sample ID", "Kit Number", "Sample Type",
  "Isolation Date", "RIN", "DV200 (%)"
)

annotSampleEM15 <- annotSampleEM15[, c(
  "RNA Sample ID", "DECAMP #",
  "Sample ID", "Kit Number", "Sample Type",
  "Isolation Date", "RIN", "DV200 (%)"
)]
colnames(annotSampleEM15) <- c(
  "RNA Sample ID", "DECAMP #",
  "Sample ID", "Kit Number", "Sample Type",
  "Isolation Date", "RIN", "DV200 (%)"
)

annotSampleEM16 <- annotSampleEM16[, c(
  "RNA Sample ID", "DECAMP #",
  "Sample ID", "Kit Number", "Sample Type",
  "Isolation Date", "RIN", "DV200 (%)"
)]
colnames(annotSampleEM16) <- c(
  "RNA Sample ID", "DECAMP #",
  "Sample ID", "Kit Number", "Sample Type",
  "Isolation Date", "RIN", "DV200 (%)"
)

annotSampleN1 <- annotSampleN1[, c(
  "RNA Sample ID", "DECAMP #",
  "Sample ID", "Kit Number", "Sample Type",
  "Isolation Date", "RIN", "DV200 (%)"
)]
colnames(annotSampleN1) <- c(
  "RNA Sample ID", "DECAMP #",
  "Sample ID", "Kit Number", "Sample Type",
  "Isolation Date", "RIN", "DV200 (%)"
)

annotSampleN2 <- annotSampleN2[, c(
  "RNA Sample ID", "DECAMP #",
  "Sample ID", "Kit Number", "Sample Type",
  "Isolation Date", "RIN", "DV200 (%)"
)]
colnames(annotSampleN2) <- c(
  "RNA Sample ID", "DECAMP #",
  "Sample ID", "Kit Number", "Sample Type",
  "Isolation Date", "RIN", "DV200 (%)"
)

allAnnotSamples <- ls()[grep("annotSample", ls())]
annotSample <- as.data.frame(matrix(ncol = 8))
colnames(annotSample) <- c(
  "RNA Sample ID", "DECAMP #",
  "Sample ID", "Kit Number", "Sample Type",
  "Isolation Date", "RIN", "DV200 (%)"
)
# Combine
for (x in allAnnotSamples) {
  annotSamples <- get(x)
  annotSamples$`Isolation Date` <- as.character(annotSamples$`Isolation Date`)
  annotSample <- rbind(annotSample, annotSamples)
}
annotSample <- annotSample[-which(is.na(annotSample$`RNA Sample ID`)), ]
annotSample <- annotSample[-which(is.na(annotSample$`Sample Type`)), ]

annotSample <- annotSample[-which(duplicated(annotSample$`RNA Sample ID`)), ]

# Save which samples had wrong rRNA peak when checking RIN
wrongRRNA.ix <- grep(pattern = "rong rRNA peak", annotSample$RIN)
annotSample$wrongRNAPeak <- 0
annotSample$wrongRNAPeak[wrongRRNA.ix] <- 1

# Edit RIN section to convert into numeric vector
annotSample$RIN <- gsub(pattern = "(wrong rRNA peak)", replacement = "", annotSample$RIN, fixed = TRUE)
annotSample$RIN <- gsub(pattern = "(wrong RNA peak)", replacement = "", annotSample$RIN, fixed = TRUE)
annotSample$RIN <- gsub(pattern = "(wrong rRNA peaks)", replacement = "", annotSample$RIN, fixed = TRUE)
annotSample$RIN <- gsub(pattern = "(Wrong rRNA peak)", replacement = "", annotSample$RIN, fixed = TRUE)
annotSample$RIN <- gsub(pattern = "(Wrong rRNA peaks)", replacement = "", annotSample$RIN, fixed = TRUE)
annotSample$RIN <- gsub(pattern = "\\/NA", replacement = "", annotSample$RIN)
annotSample$RIN <- gsub(pattern = ".*\\/", replacement = "", annotSample$RIN)
annotSample$RIN <- as.numeric(annotSample$RIN)
# There is a "29", probably 2.9, will ask Jack
annotSample$RIN[annotSample$RIN == 29] <- 2.9

# Consolidate nasal/bronch terms
annotSample$`Sample Type` <- gsub(pattern = ".*Bronch.*", replacement = "Bronchial Brushings", annotSample$`Sample Type`)
annotSample$`Sample Type` <- gsub(pattern = ".*Nasal.*", replacement = "Nasal Brushings", annotSample$`Sample Type`)

annotManifest <- filter(annotManifest, Run == "Run 1")
annotSample <- annotSample[annotSample$`RNA Sample ID` %in% annotManifest$Sample, ]
annotSample <- annotSample[match(annotSample$`RNA Sample ID`, annotManifest$Sample), ]

# Removing duplicate column
annotManifest$`Kit Number` <- NULL

annot <- cbind(annotManifest, annotSample)

# Clinical info:

# Matches kitnumber to randID, "Novartis_datadump"
annotDataDump2 <- read_excel("/rprojectnb/decamp/annotation/20200920/FullDECAMP2_DataDump_20200920.xlsx", sheet = "_D2CASEMAP", )

# randID to demographics info, "Novartis_datadump"
annotDataDump3 <- read_excel("/rprojectnb/decamp/annotation/20200920/FullDECAMP2_DataDump_20200920.xlsx", sheet = "_D2DEMO")

# randID to age/smoking info, "Novartis_datadump"
annotDataDump4 <- read_excel("/rprojectnb/decamp/annotation/20200920/FullDECAMP2_DataDump_20200920.xlsx", sheet = "_D2ELIGCHECK")

# randID to FEV/FVC, "Novartis_datadump"
annotDataDump5 <- read_excel("/rprojectnb/decamp/annotation/20200920/FullDECAMP2_DataDump_20200920.xlsx", sheet = "_D2PFT")
annotDataDump5 <- filter(annotDataDump5, timepoint == "Baseline Visit")

# Combine annotation info
annotA <- merge(merge(annotDataDump2, annotDataDump3, by.x = "randID", by.y = "randID"), annotDataDump4, by.x = "randID", by.y = "randID")
annotA <- merge(annotA, annotDataDump5, by.x = "randID", by.y = "randID")

# Removing patient "4796-053" as "Pacific Islander" for time being
annotA <- annotA[!(duplicated(annotA$randID)), ]

annot <- merge(annot, annotA, by.x = "Kit Number", by.y = "kitnumber", all.x = TRUE)

annotBronch <- annot[annot$`Sample Type` == "Bronchial Brushings", ]
annotNasal <- annot[annot$`Sample Type` == "Nasal Brushings", ]

uniquePatients <- length(unique(annot$`Kit Number`))

numBronch <- length(unique(annotBronch$Sample))
numNasal <- length(unique(annotNasal$Sample))

bothBronchNasal <- length(intersect(annotBronch$`Kit Number`, annotNasal$`Kit Number`))
```
There are `r length(annot$Sample)` samples and `r uniquePatients` unique patients in the dataset. `r numBronch` are bronchial samples, and `r numNasal` are nasal samples. There are `r bothBronchNasal` patients that have both bronchial and nasal samples.


```{r}
# Samples flagged by Novartis as poor quality:
flaggedByNovartis <- c(
  "EM15-14", "EM15-31", "EM15-64", "EM15-85", "EM16-194", "EM16-195",
  "EM16-228", "EM5-100", "EM8-112", "EM8-130", "EM8-158", "EM8-171",
  "EM8-26", "EM8-55", "EM8-66", "N1-115", "N1-119", "N1-125",
  "N1-139", "N1-25", "N1-40", "N1-55", "N1-63", "N1-64",
  "N1-67", "N1-81", "N1-83", "N1-84", "N1-94", "N2-10",
  "N2-15", "N2-18", "N2-30", "N2-31", "N2-22", "N2-28"
)
```

```{r, warning = F, message = F}
seBronch <- se[, which(colnames(se) %in% annotBronch$`RNA Sample ID`)]

annotBronch <- annotBronch[annotBronch$`RNA Sample ID` %in% colnames(se), ]

# reorder
annotBronch <- annotBronch[match(colnames(seBronch), annotBronch$`RNA Sample ID`), ]

coldataBronch <- colData(seBronch)
```

## Violin Plots, QC Metrics {.tabset}

### RIN

```{r, warning = F, message = F}
dots <- T
boxplot <- F
violin <- T

df <- data.frame(
  x = annotBronch$`Sample Type`,
  y = annotBronch$RIN
)

p <- ggplot2::ggplot(df) +
  ggplot2::aes_string(
    x = "x",
    y = "y"
  )
if (dots == TRUE) {
  p <- p + ggplot2::geom_jitter(
    color = "blue",
    width = 0.2,
    height = 0,
    size = 1,
    alpha = 1
  )
}
if (boxplot == TRUE) {
  p <- p + ggplot2::geom_boxplot(
    width = 0.5,
    alpha = 0
  )
}
if (violin == TRUE) {
  p <- p + ggplot2::geom_violin(
    trim = TRUE,
    scale = "width",
    size = 1,
    fill = "grey",
    alpha = 0.75
  )
}

p <- p + ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 14),
    axis.title = ggplot2::element_text(size = 10)
  ) + ggplot2::theme(axis.text.x = ggplot2::element_text(size = 15)) +
  ggplot2::ylab("RIN") +
  ggplot2::theme(
    axis.title.y = element_text(size = 15),
    axis.title.x = ggplot2::element_blank()
  )

p <- p + ggplot2::ggtitle(label = "Bronchial Brushings") +
  ggplot2::theme(plot.title = ggplot2::element_text(
    hjust = 0.5,
    size = 18
  ))

p
```

### Genes detected

```{r, warning = F, message = F}
dots <- T
boxplot <- T
violin <- T

df <- data.frame(
  x = annotBronch$`Sample Type`,
  y = colSums(assay(seBronch) > 0)
)

p <- ggplot2::ggplot(df) +
  ggplot2::aes_string(
    x = "x",
    y = "y"
  )
if (dots == TRUE) {
  p <- p + ggplot2::geom_jitter(
    color = "blue",
    width = 0.2,
    height = 0,
    size = 1,
    alpha = 1
  )
}
if (boxplot == TRUE) {
  p <- p + ggplot2::geom_boxplot(
    width = 0.5,
    alpha = 0
  )
}
if (violin == TRUE) {
  p <- p + ggplot2::geom_violin(
    trim = TRUE,
    scale = "width",
    size = 1,
    fill = "grey",
    alpha = 0.75
  )
}

p <- p + ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 10),
    axis.title = ggplot2::element_text(size = 10)
  ) + ggplot2::theme(axis.text.x = ggplot2::element_text(size = 15)) +
  ggplot2::ylab("# Genes detected") +
  ggplot2::theme(
    axis.title.y = element_text(size = 15),
    axis.title.x = ggplot2::element_blank()
  )

p <- p + ggplot2::ggtitle(label = "Number of genes detected") +
  ggplot2::theme(plot.title = ggplot2::element_text(
    hjust = 0.5,
    size = 18
  ))

summary <- "median"
summ <- df %>%
  dplyr::group_by(x) %>%
  dplyr::summarize(value = stats::median(y))
fun <- stats::median

summ$statY <- max(df$y) + (max(df$y) - min(df$y)) * 0.1
summary <- paste(toupper(substr(summary, 1, 1)),
  substr(summary, 2, nchar(summary)),
  sep = ""
)
summ$label <- paste0(summary, ": ", round(summ$value, 5))

p <- p + ggplot2::geom_text(
  data = summ,
  ggplot2::aes_string(
    x = "x",
    y = "statY",
    label = "label"
  ),
  size = 5
)
p <- p + ggplot2::stat_summary(
  fun = fun, fun.min = fun,
  fun.max = fun,
  geom = "crossbar",
  color = "red",
  linetype = "dashed"
)

p
```

### Percent GC

```{r, warning = F, message = F}

df <- data.frame(
  x = annotBronch$`Sample Type`,
  y = unlist(lapply(coldataBronch$FastQC_percent_GC, FUN = mean))
)

p <- ggplot2::ggplot(df) +
  ggplot2::aes_string(
    x = "x",
    y = "y"
  )
if (dots == TRUE) {
  p <- p + ggplot2::geom_jitter(
    color = "blue",
    width = 0.2,
    height = 0,
    size = 1,
    alpha = 1
  )
}
if (boxplot == TRUE) {
  p <- p + ggplot2::geom_boxplot(
    width = 0.5,
    alpha = 0
  )
}
if (violin == TRUE) {
  p <- p + ggplot2::geom_violin(
    trim = TRUE,
    scale = "width",
    size = 1,
    fill = "grey",
    alpha = 0.75
  )
}

p <- p + ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 10),
    axis.title = ggplot2::element_text(size = 10)
  ) + ggplot2::theme(axis.text.x = ggplot2::element_text(size = 15)) +
  ggplot2::ylab("Mean GC content (%)") +
  ggplot2::theme(
    axis.title.y = element_text(size = 15),
    axis.title.x = ggplot2::element_blank()
  )

p <- p + ggplot2::ggtitle(label = "FastQC, GC content (%)") +
  ggplot2::theme(plot.title = ggplot2::element_text(
    hjust = 0.5,
    size = 18
  ))

summary <- "median"
summ <- df %>%
  dplyr::group_by(x) %>%
  dplyr::summarize(value = stats::median(y))
fun <- stats::median

summ$statY <- max(df$y) + (max(df$y) - min(df$y)) * 0.1
summary <- paste(toupper(substr(summary, 1, 1)),
  substr(summary, 2, nchar(summary)),
  sep = ""
)
summ$label <- paste0(summary, ": ", round(summ$value, 5))

p <- p + ggplot2::geom_text(
  data = summ,
  ggplot2::aes_string(
    x = "x",
    y = "statY",
    label = "label"
  ),
  size = 5
)
p <- p + ggplot2::stat_summary(
  fun = fun, fun.min = fun,
  fun.max = fun,
  geom = "crossbar",
  color = "red",
  linetype = "dashed"
)

p
```

### Percentage aligned to exons

```{r, warning = F, message = F}
df <- data.frame(
  x = annotBronch$`Sample Type`,
  y = (coldataBronch$RSeQC_cds_exons_tag_pct +
    coldataBronch$RSeQC_3_utr_exons_tag_pct +
    coldataBronch$RSeQC_5_utr_exons_tag_pct)
)

p <- ggplot2::ggplot(df) +
  ggplot2::aes_string(
    x = "x",
    y = "y"
  )
if (dots == TRUE) {
  p <- p + ggplot2::geom_jitter(
    color = "blue",
    width = 0.2,
    height = 0,
    size = 1,
    alpha = 1
  )
}
if (boxplot == TRUE) {
  p <- p + ggplot2::geom_boxplot(
    width = 0.5,
    alpha = 0
  )
}
if (violin == TRUE) {
  p <- p + ggplot2::geom_violin(
    trim = TRUE,
    scale = "width",
    size = 1,
    fill = "grey",
    alpha = 0.75
  )
}

p <- p + ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 10),
    axis.title = ggplot2::element_text(size = 10)
  ) + ggplot2::theme(axis.text.x = ggplot2::element_text(size = 15)) +
  ggplot2::ylab("Aligned to exonic region (%)") +
  ggplot2::theme(
    axis.title.y = element_text(size = 15),
    axis.title.x = ggplot2::element_blank()
  )

p <- p + ggplot2::ggtitle(label = "RSeQC, Percent aligned to exons") +
  ggplot2::theme(plot.title = ggplot2::element_text(
    hjust = 0.5,
    size = 18
  ))

summary <- "median"
summ <- df %>%
  dplyr::group_by(x) %>%
  dplyr::summarize(value = stats::median(y))
fun <- stats::median

summ$statY <- max(df$y) + (max(df$y) - min(df$y)) * 0.1
summary <- paste(toupper(substr(summary, 1, 1)),
  substr(summary, 2, nchar(summary)),
  sep = ""
)
summ$label <- paste0(summary, ": ", round(summ$value, 5))

p <- p + ggplot2::geom_text(
  data = summ,
  ggplot2::aes_string(
    x = "x",
    y = "statY",
    label = "label"
  ),
  size = 5
)
p <- p + ggplot2::stat_summary(
  fun = fun, fun.min = fun,
  fun.max = fun,
  geom = "crossbar",
  color = "red",
  linetype = "dashed"
)

p
```

### Percentage aligned to introns

```{r, warning = F, message = F}
df <- data.frame(
  x = annotBronch$`Sample Type`,
  y = coldataBronch$RSeQC_introns_tag_pct
)

p <- ggplot2::ggplot(df) +
  ggplot2::aes_string(
    x = "x",
    y = "y"
  )
if (dots == TRUE) {
  p <- p + ggplot2::geom_jitter(
    color = "blue",
    width = 0.2,
    height = 0,
    size = 1,
    alpha = 1
  )
}
if (boxplot == TRUE) {
  p <- p + ggplot2::geom_boxplot(
    width = 0.5,
    alpha = 0
  )
}
if (violin == TRUE) {
  p <- p + ggplot2::geom_violin(
    trim = TRUE,
    scale = "width",
    size = 1,
    fill = "grey",
    alpha = 0.75
  )
}

p <- p + ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 10),
    axis.title = ggplot2::element_text(size = 10)
  ) + ggplot2::theme(axis.text.x = ggplot2::element_text(size = 15)) +
  ggplot2::ylab("Aligned to intronic region (%)") +
  ggplot2::theme(
    axis.title.y = element_text(size = 15),
    axis.title.x = ggplot2::element_blank()
  )

p <- p + ggplot2::ggtitle(label = "RSeQC, Percent aligned to introns") +
  ggplot2::theme(plot.title = ggplot2::element_text(
    hjust = 0.5,
    size = 18
  ))

summary <- "median"
summ <- df %>%
  dplyr::group_by(x) %>%
  dplyr::summarize(value = stats::median(y))
fun <- stats::median

summ$statY <- max(df$y) + (max(df$y) - min(df$y)) * 0.1
summary <- paste(toupper(substr(summary, 1, 1)),
  substr(summary, 2, nchar(summary)),
  sep = ""
)
summ$label <- paste0(summary, ": ", round(summ$value, 5))

p <- p + ggplot2::geom_text(
  data = summ,
  ggplot2::aes_string(
    x = "x",
    y = "statY",
    label = "label"
  ),
  size = 5
)
p <- p + ggplot2::stat_summary(
  fun = fun, fun.min = fun,
  fun.max = fun,
  geom = "crossbar",
  color = "red",
  linetype = "dashed"
)

p
```

### Percentage aligned to intergenic regions

```{r, warning = F, message = F}
df <- data.frame(
  x = annotBronch$`Sample Type`,
  y = coldataBronch$RSeQC_tes_down_10kb_tag_pct +
    coldataBronch$RSeQC_tss_up_10kb_tag_pct
)

p <- ggplot2::ggplot(df) +
  ggplot2::aes_string(
    x = "x",
    y = "y"
  )
if (dots == TRUE) {
  p <- p + ggplot2::geom_jitter(
    color = "blue",
    width = 0.2,
    height = 0,
    size = 1,
    alpha = 1
  )
}
if (boxplot == TRUE) {
  p <- p + ggplot2::geom_boxplot(
    width = 0.5,
    alpha = 0
  )
}
if (violin == TRUE) {
  p <- p + ggplot2::geom_violin(
    trim = TRUE,
    scale = "width",
    size = 1,
    fill = "grey",
    alpha = 0.75
  )
}

p <- p + ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 10),
    axis.title = ggplot2::element_text(size = 10)
  ) + ggplot2::theme(axis.text.x = ggplot2::element_text(size = 15)) +
  ggplot2::ylab("Aligned to intergenic region (%)") +
  ggplot2::theme(
    axis.title.y = element_text(size = 15),
    axis.title.x = ggplot2::element_blank()
  )

p <- p + ggplot2::ggtitle(label = "RSeQC, Percent aligned to intergenic region") +
  ggplot2::theme(plot.title = ggplot2::element_text(
    hjust = 0.5,
    size = 18
  ))

summary <- "median"
summ <- df %>%
  dplyr::group_by(x) %>%
  dplyr::summarize(value = stats::median(y))
fun <- stats::median

summ$statY <- max(df$y) + (max(df$y) - min(df$y)) * 0.1
summary <- paste(toupper(substr(summary, 1, 1)),
  substr(summary, 2, nchar(summary)),
  sep = ""
)
summ$label <- paste0(summary, ": ", round(summ$value, 5))

p <- p + ggplot2::geom_text(
  data = summ,
  ggplot2::aes_string(
    x = "x",
    y = "statY",
    label = "label"
  ),
  size = 5
)
p <- p + ggplot2::stat_summary(
  fun = fun, fun.min = fun,
  fun.max = fun,
  geom = "crossbar",
  color = "red",
  linetype = "dashed"
)

p
```
Includes regions upstream of TSS (<10kb) and downstream of TES (<10kb)

### Strandedness - Sense

```{r, warning = F, message = F}
df <- data.frame(
  x = annotBronch$`Sample Type`,
  y = coldataBronch$RSeQC_pe_sense * 100
)

p <- ggplot2::ggplot(df) +
  ggplot2::aes_string(
    x = "x",
    y = "y"
  )
if (dots == TRUE) {
  p <- p + ggplot2::geom_jitter(
    color = "blue",
    width = 0.2,
    height = 0,
    size = 1,
    alpha = 1
  )
}
if (boxplot == TRUE) {
  p <- p + ggplot2::geom_boxplot(
    width = 0.5,
    alpha = 0
  )
}
if (violin == TRUE) {
  p <- p + ggplot2::geom_violin(
    trim = TRUE,
    scale = "width",
    size = 1,
    fill = "grey",
    alpha = 0.75
  )
}

p <- p + ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 10),
    axis.title = ggplot2::element_text(size = 10)
  ) + ggplot2::theme(axis.text.x = ggplot2::element_text(size = 15)) +
  ggplot2::ylab("Sense-stranded (%)") +
  ggplot2::theme(
    axis.title.y = element_text(size = 15),
    axis.title.x = ggplot2::element_blank()
  )

p <- p + ggplot2::ggtitle(label = "RSeQC, Percent Sense-strandedness") +
  ggplot2::theme(plot.title = ggplot2::element_text(
    hjust = 0.5,
    size = 18
  ))

summary <- "median"
summ <- df %>%
  dplyr::group_by(x) %>%
  dplyr::summarize(value = stats::median(y))
fun <- stats::median

summ$statY <- max(df$y) + (max(df$y) - min(df$y)) * 0.1
summary <- paste(toupper(substr(summary, 1, 1)),
  substr(summary, 2, nchar(summary)),
  sep = ""
)
summ$label <- paste0(summary, ": ", round(summ$value, 5))

p <- p + ggplot2::geom_text(
  data = summ,
  ggplot2::aes_string(
    x = "x",
    y = "statY",
    label = "label"
  ),
  size = 5
)
p <- p + ggplot2::stat_summary(
  fun = fun, fun.min = fun,
  fun.max = fun,
  geom = "crossbar",
  color = "red",
  linetype = "dashed"
)

p
```

### Strandedness - Antisense

```{r, warning = F, message = F}
df <- data.frame(
  x = annotBronch$`Sample Type`,
  y = coldataBronch$RSeQC_pe_antisense * 100
)

p <- ggplot2::ggplot(df) +
  ggplot2::aes_string(
    x = "x",
    y = "y"
  )
if (dots == TRUE) {
  p <- p + ggplot2::geom_jitter(
    color = "blue",
    width = 0.2,
    height = 0,
    size = 1,
    alpha = 1
  )
}
if (boxplot == TRUE) {
  p <- p + ggplot2::geom_boxplot(
    width = 0.5,
    alpha = 0
  )
}
if (violin == TRUE) {
  p <- p + ggplot2::geom_violin(
    trim = TRUE,
    scale = "width",
    size = 1,
    fill = "grey",
    alpha = 0.75
  )
}

p <- p + ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 10),
    axis.title = ggplot2::element_text(size = 10)
  ) + ggplot2::theme(axis.text.x = ggplot2::element_text(size = 15)) +
  ggplot2::ylab("Antisense-stranded (%)") +
  ggplot2::theme(
    axis.title.y = element_text(size = 15),
    axis.title.x = ggplot2::element_blank()
  )

p <- p + ggplot2::ggtitle(label = "RSeQC, Percent Antisense-strandedness") +
  ggplot2::theme(plot.title = ggplot2::element_text(
    hjust = 0.5,
    size = 18
  ))

summary <- "median"
summ <- df %>%
  dplyr::group_by(x) %>%
  dplyr::summarize(value = stats::median(y))
fun <- stats::median

summ$statY <- max(df$y) + (max(df$y) - min(df$y)) * 0.1
summary <- paste(toupper(substr(summary, 1, 1)),
  substr(summary, 2, nchar(summary)),
  sep = ""
)
summ$label <- paste0(summary, ": ", round(summ$value, 5))

p <- p + ggplot2::geom_text(
  data = summ,
  ggplot2::aes_string(
    x = "x",
    y = "statY",
    label = "label"
  ),
  size = 5
)
p <- p + ggplot2::stat_summary(
  fun = fun, fun.min = fun,
  fun.max = fun,
  geom = "crossbar",
  color = "red",
  linetype = "dashed"
)

p
```


### Proper Pair Percentage

```{r, warning = F, message = F}
df <- data.frame(
  x = annotBronch$`Sample Type`,
  y = coldataBronch$RSeQC_proper_pairs_percent
)

p <- ggplot2::ggplot(df) +
  ggplot2::aes_string(
    x = "x",
    y = "y"
  )
if (dots == TRUE) {
  p <- p + ggplot2::geom_jitter(
    color = "blue",
    width = 0.2,
    height = 0,
    size = 1,
    alpha = 1
  )
}
if (boxplot == TRUE) {
  p <- p + ggplot2::geom_boxplot(
    width = 0.5,
    alpha = 0
  )
}
if (violin == TRUE) {
  p <- p + ggplot2::geom_violin(
    trim = TRUE,
    scale = "width",
    size = 1,
    fill = "grey",
    alpha = 0.75
  )
}

p <- p + ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 10),
    axis.title = ggplot2::element_text(size = 10)
  ) + ggplot2::theme(axis.text.x = ggplot2::element_text(size = 15)) +
  ggplot2::ylab("TIN") +
  ggplot2::theme(
    axis.title.y = element_text(size = 15),
    axis.title.x = ggplot2::element_blank()
  )

p <- p + ggplot2::ggtitle(label = "RSeQC, Proper Pairs (%)") +
  ggplot2::theme(plot.title = ggplot2::element_text(
    hjust = 0.5,
    size = 18
  ))
summary <- "median"
summ <- df %>%
  dplyr::group_by(x) %>%
  dplyr::summarize(value = stats::median(y))
fun <- stats::median

summ$statY <- max(df$y) + (max(df$y) - min(df$y)) * 0.1
summary <- paste(toupper(substr(summary, 1, 1)),
  substr(summary, 2, nchar(summary)),
  sep = ""
)
summ$label <- paste0(summary, ": ", round(summ$value, 5))

p <- p + ggplot2::geom_text(
  data = summ,
  ggplot2::aes_string(
    x = "x",
    y = "statY",
    label = "label"
  ),
  size = 5
)
p <- p + ggplot2::stat_summary(
  fun = fun, fun.min = fun,
  fun.max = fun,
  geom = "crossbar",
  color = "red",
  linetype = "dashed"
)
p
```

### TIN

```{r, warning = F, message = F}
df <- data.frame(
  x = annotBronch$`Sample Type`,
  y = coldataBronch$RSeQC_TIN_mean
)

p <- ggplot2::ggplot(df) +
  ggplot2::aes_string(
    x = "x",
    y = "y"
  )
if (dots == TRUE) {
  p <- p + ggplot2::geom_jitter(
    color = "blue",
    width = 0.2,
    height = 0,
    size = 1,
    alpha = 1
  )
}
if (boxplot == TRUE) {
  p <- p + ggplot2::geom_boxplot(
    width = 0.5,
    alpha = 0
  )
}
if (violin == TRUE) {
  p <- p + ggplot2::geom_violin(
    trim = TRUE,
    scale = "width",
    size = 1,
    fill = "grey",
    alpha = 0.75
  )
}

p <- p + ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 10),
    axis.title = ggplot2::element_text(size = 10)
  ) + ggplot2::theme(axis.text.x = ggplot2::element_text(size = 15)) +
  ggplot2::ylab("TIN") +
  ggplot2::theme(
    axis.title.y = element_text(size = 15),
    axis.title.x = ggplot2::element_blank()
  )

p <- p + ggplot2::ggtitle(label = "RSeQC, TIN") +
  ggplot2::theme(plot.title = ggplot2::element_text(
    hjust = 0.5,
    size = 18
  ))

summary <- "median"
summ <- df %>%
  dplyr::group_by(x) %>%
  dplyr::summarize(value = stats::median(y))
fun <- stats::median

summ$statY <- max(df$y) + (max(df$y) - min(df$y)) * 0.1
summary <- paste(toupper(substr(summary, 1, 1)),
  substr(summary, 2, nchar(summary)),
  sep = ""
)
summ$label <- paste0(summary, ": ", round(summ$value, 5))

p <- p + ggplot2::geom_text(
  data = summ,
  ggplot2::aes_string(
    x = "x",
    y = "statY",
    label = "label"
  ),
  size = 5
)
p <- p + ggplot2::stat_summary(
  fun = fun, fun.min = fun,
  fun.max = fun,
  geom = "crossbar",
  color = "red",
  linetype = "dashed"
)

p
```


### Percentage reads uniquely mapped

```{r, warning = F, message = F}
df <- data.frame(
  x = annotBronch$`Sample Type`,
  y = coldataBronch$STAR_uniquely_mapped_percent
)

p <- ggplot2::ggplot(df) +
  ggplot2::aes_string(
    x = "x",
    y = "y"
  )
if (dots == TRUE) {
  p <- p + ggplot2::geom_jitter(
    color = "blue",
    width = 0.2,
    height = 0,
    size = 1,
    alpha = 1
  )
}
if (boxplot == TRUE) {
  p <- p + ggplot2::geom_boxplot(
    width = 0.5,
    alpha = 0
  )
}
if (violin == TRUE) {
  p <- p + ggplot2::geom_violin(
    trim = TRUE,
    scale = "width",
    size = 1,
    fill = "grey",
    alpha = 0.75
  )
}

p <- p + ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 10),
    axis.title = ggplot2::element_text(size = 10)
  ) + ggplot2::theme(axis.text.x = ggplot2::element_text(size = 15)) +
  ggplot2::ylab("Uniquely mapped to genome(%)") +
  ggplot2::theme(
    axis.title.y = element_text(size = 15),
    axis.title.x = ggplot2::element_blank()
  )

p <- p + ggplot2::ggtitle(label = "STAR, Unique mapped") +
  ggplot2::theme(plot.title = ggplot2::element_text(
    hjust = 0.5,
    size = 18
  ))
summary <- "median"
summ <- df %>%
  dplyr::group_by(x) %>%
  dplyr::summarize(value = stats::median(y))
fun <- stats::median

summ$statY <- max(df$y) + (max(df$y) - min(df$y)) * 0.1
summary <- paste(toupper(substr(summary, 1, 1)),
  substr(summary, 2, nchar(summary)),
  sep = ""
)
summ$label <- paste0(summary, ": ", round(summ$value, 5))

p <- p + ggplot2::geom_text(
  data = summ,
  ggplot2::aes_string(
    x = "x",
    y = "statY",
    label = "label"
  ),
  size = 5
)
p <- p + ggplot2::stat_summary(
  fun = fun, fun.min = fun,
  fun.max = fun,
  geom = "crossbar",
  color = "red",
  linetype = "dashed"
)

p
```


### Percentage reads unmapped

```{r, warning = F, message = F}
df <- data.frame(
  x = annotBronch$`Sample Type`,
  y = coldataBronch$STAR_unmapped_mismatches_percent +
    coldataBronch$STAR_unmapped_tooshort_percent +
    coldataBronch$STAR_unmapped_other_percent
)

p <- ggplot2::ggplot(df) +
  ggplot2::aes_string(
    x = "x",
    y = "y"
  )
if (dots == TRUE) {
  p <- p + ggplot2::geom_jitter(
    color = "blue",
    width = 0.2,
    height = 0,
    size = 1,
    alpha = 1
  )
}
if (boxplot == TRUE) {
  p <- p + ggplot2::geom_boxplot(
    width = 0.5,
    alpha = 0
  )
}
if (violin == TRUE) {
  p <- p + ggplot2::geom_violin(
    trim = TRUE,
    scale = "width",
    size = 1,
    fill = "grey",
    alpha = 0.75
  )
}

p <- p + ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 10),
    axis.title = ggplot2::element_text(size = 10)
  ) + ggplot2::theme(axis.text.x = ggplot2::element_text(size = 15)) +
  ggplot2::ylab("Unmapped to genome(%)") +
  ggplot2::theme(
    axis.title.y = element_text(size = 15),
    axis.title.x = ggplot2::element_blank()
  )

p <- p + ggplot2::ggtitle(label = "STAR, Percent unmapped") +
  ggplot2::theme(plot.title = ggplot2::element_text(
    hjust = 0.5,
    size = 18
  ))
summary <- "median"
summ <- df %>%
  dplyr::group_by(x) %>%
  dplyr::summarize(value = stats::median(y))
fun <- stats::median

summ$statY <- max(df$y) + (max(df$y) - min(df$y)) * 0.1
summary <- paste(toupper(substr(summary, 1, 1)),
  substr(summary, 2, nchar(summary)),
  sep = ""
)
summ$label <- paste0(summary, ": ", round(summ$value, 5))

p <- p + ggplot2::geom_text(
  data = summ,
  ggplot2::aes_string(
    x = "x",
    y = "statY",
    label = "label"
  ),
  size = 5
)
p <- p + ggplot2::stat_summary(
  fun = fun, fun.min = fun,
  fun.max = fun,
  geom = "crossbar",
  color = "red",
  linetype = "dashed"
)
p
```

### Total reads

```{r, warning = F, message = F}
df <- data.frame(
  x = annotBronch$`Sample Type`,
  y = coldataBronch$STAR_total_reads
)

p <- ggplot2::ggplot(df) +
  ggplot2::aes_string(
    x = "x",
    y = "y"
  )
if (dots == TRUE) {
  p <- p + ggplot2::geom_jitter(
    color = "blue",
    width = 0.2,
    height = 0,
    size = 1,
    alpha = 1
  )
}
if (boxplot == TRUE) {
  p <- p + ggplot2::geom_boxplot(
    width = 0.5,
    alpha = 0
  )
}
if (violin == TRUE) {
  p <- p + ggplot2::geom_violin(
    trim = TRUE,
    scale = "width",
    size = 1,
    fill = "grey",
    alpha = 0.75
  )
}

p <- p + ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 10),
    axis.title = ggplot2::element_text(size = 10)
  ) + ggplot2::theme(axis.text.x = ggplot2::element_text(size = 15)) +
  ggplot2::ylab("Total Number of Reads") +
  ggplot2::theme(
    axis.title.y = element_text(size = 15),
    axis.title.x = ggplot2::element_blank()
  )

p <- p + ggplot2::ggtitle(label = "STAR, Total reads") +
  ggplot2::theme(plot.title = ggplot2::element_text(
    hjust = 0.5,
    size = 18
  ))
summary <- "median"
summ <- df %>%
  dplyr::group_by(x) %>%
  dplyr::summarize(value = stats::median(y))
fun <- stats::median

summ$statY <- max(df$y) + (max(df$y) - min(df$y)) * 0.1
summary <- paste(toupper(substr(summary, 1, 1)),
  substr(summary, 2, nchar(summary)),
  sep = ""
)
summ$label <- paste0(summary, ": ", round(summ$value, 5))

p <- p + ggplot2::geom_text(
  data = summ,
  ggplot2::aes_string(
    x = "x",
    y = "statY",
    label = "label"
  ),
  size = 5
)
p <- p + ggplot2::stat_summary(
  fun = fun, fun.min = fun,
  fun.max = fun,
  geom = "crossbar",
  color = "red",
  linetype = "dashed"
)
p
```


### Mismatch rate

```{r, warning = F, message = F}
df <- data.frame(
  x = annotBronch$`Sample Type`,
  y = coldataBronch$STAR_mismatch_rate
)

p <- ggplot2::ggplot(df) +
  ggplot2::aes_string(
    x = "x",
    y = "y"
  )
if (dots == TRUE) {
  p <- p + ggplot2::geom_jitter(
    color = "blue",
    width = 0.2,
    height = 0,
    size = 1,
    alpha = 1
  )
}
if (boxplot == TRUE) {
  p <- p + ggplot2::geom_boxplot(
    width = 0.5,
    alpha = 0
  )
}
if (violin == TRUE) {
  p <- p + ggplot2::geom_violin(
    trim = TRUE,
    scale = "width",
    size = 1,
    fill = "grey",
    alpha = 0.75
  )
}

p <- p + ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 10),
    axis.title = ggplot2::element_text(size = 10)
  ) + ggplot2::theme(axis.text.x = ggplot2::element_text(size = 15)) +
  ggplot2::ylab("Mismatch rate") +
  ggplot2::theme(
    axis.title.y = element_text(size = 15),
    axis.title.x = ggplot2::element_blank()
  )

p <- p + ggplot2::ggtitle(label = "STAR, mismatch rate") +
  ggplot2::theme(plot.title = ggplot2::element_text(
    hjust = 0.5,
    size = 18
  ))
summary <- "median"
summ <- df %>%
  dplyr::group_by(x) %>%
  dplyr::summarize(value = stats::median(y))
fun <- stats::median

summ$statY <- max(df$y) + (max(df$y) - min(df$y)) * 0.1
summary <- paste(toupper(substr(summary, 1, 1)),
  substr(summary, 2, nchar(summary)),
  sep = ""
)
summ$label <- paste0(summary, ": ", round(summ$value, 5))

p <- p + ggplot2::geom_text(
  data = summ,
  ggplot2::aes_string(
    x = "x",
    y = "statY",
    label = "label"
  ),
  size = 5
)
p <- p + ggplot2::stat_summary(
  fun = fun, fun.min = fun,
  fun.max = fun,
  geom = "crossbar",
  color = "red",
  linetype = "dashed"
)
p
```

## Histogram, QC outliers

A tally of how many times a sample had an outlier for a numeric QC output was taken. An outlier was considered to be a value past 2 standard deviations of the mean. The barplot below is colored by whether the sample RIN was higher or lower than 5, to determine if RIN scores had a correlation to the QC results.

```{r, message = FALSE, warning=FALSE}
fastqc.ix <- grep("FastQC", colnames(coldataBronch))
rseqc.ix <- grep("RSeQC", colnames(coldataBronch))
star.ix <- grep("STAR", colnames(coldataBronch))

colDataBronchSubset <- coldataBronch[, c(fastqc.ix, rseqc.ix, star.ix)]

# Bronch samples flagged by Novartis
flaggedByNovartisBronch <- flaggedByNovartis[flaggedByNovartis %in% rownames(colDataBronchSubset)]
novartisQC <- sapply(rownames(colDataBronchSubset), function(x) {
  if (x %in% flaggedByNovartisBronch) {
    return("Poor")
  } else {
    return("Good")
  }
})

# FastQC outputs lists instead of numeric/integer, because it is looking
# at pairs of fastqs.
classlist <- c()
for (x in 1:ncol(colDataBronchSubset)) {
  classlist <- c(classlist, class(colDataBronchSubset[1, x]))
}

# Of FastQC outputs, only taking out average seq length, total sequences,
# dedup percentage, percentGC. Other metrics seem to be passing for all data points
colDataBronchSubsetFastq <- colDataBronchSubset[, c(4, 12, 16, 17)]
colDataBronchSubsetFastq2 <- apply(colDataBronchSubsetFastq, 1:2, function(x) {
  return(mean(unlist(x)))
})
colDataBronchSubset <- colDataBronchSubset[, -which(classlist == "list")]
colDataBronchSubset <- cbind(colDataBronchSubset, colDataBronchSubsetFastq2)

# Also removing all QC metrics which have the same values across all samples
rm.ix.qc <- c()
for (x in 1:ncol(colDataBronchSubset)) {
  if (length(unique(colDataBronchSubset[, x])) == 1) {
    rm.ix.qc <- c(rm.ix.qc, x)
  }
}

colDataBronchSubset <- colDataBronchSubset[, -rm.ix.qc]

qc.tally <- matrix(ncol = nrow(colDataBronchSubset), data = 0)
colnames(qc.tally) <- rownames(colDataBronchSubset)

qc.tally2 <- matrix(
  ncol = nrow(colDataBronchSubset),
  nrow = ncol(colDataBronchSubset),
  data = 0
)

rownames(qc.tally2) <- colnames(colDataBronchSubset)
colnames(qc.tally2) <- rownames(colDataBronchSubset)

isNumeric <- 0
for (i in 1:ncol(colDataBronchSubset)) {
  qc <- colDataBronchSubset[, i]
  if (class(qc) == "numeric" | class(qc) == "integer") {
    isNumeric <- isNumeric + 1
    qc.filt <- qc[!is.na(qc)]
    mean.qc <- mean(qc.filt)
    sd.qc <- sd(qc.filt)
    qc.tally[, which(qc > mean.qc + 2 * sd.qc | qc < mean.qc - 2 * sd.qc)] <- qc.tally[, which(qc > mean.qc + 2 * sd.qc | qc < mean.qc - 2 * sd.qc)] + 1

    qc.tally2[i, which(qc > mean.qc + 2 * sd.qc | qc < mean.qc - 2 * sd.qc)] <- qc.tally2[i, which(qc > mean.qc + 2 * sd.qc | qc < mean.qc - 2 * sd.qc)] + 1
  }
}

qc.tally.gg <- as.data.frame(t(qc.tally))
qc.tally.gg$RIN <- annotBronch$RIN < 5
qc.tally.gg$RIN[qc.tally.gg$RIN == TRUE] <- "Lower than 5"
qc.tally.gg$RIN[qc.tally.gg$RIN == FALSE] <- "Higher than 5"

histQC <- ggplot(data = qc.tally.gg, aes(x = V1, color = RIN, fill = RIN)) +
  geom_histogram() +
  labs(x = "Tally of outliers") +
  ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 15),
    axis.title = ggplot2::element_text(size = 15)
  )

histQC

# rowSums(qc.tally2[,colSums(qc.tally2) > 30])

qc.tally.gg$Number_Outliers_Binned <- qc.tally.gg$V1 > 25

qc.tally.gg$Number_Outliers_Binned <- as.character(qc.tally.gg$Number_Outliers_Binned)
qc.tally.gg$Number_Outliers_Binned <- gsub("FALSE", "x < 25",
                                           qc.tally.gg$Number_Outliers_Binned)
qc.tally.gg$Number_Outliers_Binned <- gsub("TRUE", "x => 25",
                                           qc.tally.gg$Number_Outliers_Binned)

highAmountOutlier <- colnames(qc.tally)[which(qc.tally > 25)]

qualAnnot <- sapply(qc.tally, function(x) {
  if (x < 25) {
    return("x =< 25")
  } else if (x < 40) {
    return("25 < x < 40")
  } else {
    return("x >= 40")
  }
})

Number_Outliers_Binned <- as.factor(qualAnnot)
Number_Outliers_Binned <- factor(Number_Outliers_Binned, levels = c("x =< 25",
                                    "25 < x < 40",
                                    "x >= 40"))
```

A total of `r isNumeric` QC metrics were taken into consideration. The following samples had a high number of QC metrics outliers: `r highAmountOutlier`

The barplots were also annotated by the QC distinctions Novartis had previously made. (Poor vs Good)

```{r, message = FALSE, warning=FALSE}
qc.tally.gg$NovartisQC <- novartisQC
histQC2 <- ggplot(data = qc.tally.gg, aes(x = V1, color = NovartisQC, fill = NovartisQC)) +
  geom_histogram() +
  labs(x = "Tally of outliers") +
  ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 15),
    axis.title = ggplot2::element_text(size = 15)
  ) +
  scale_fill_manual(values = c("Red", "Blue"))

histQC2
```

## Using Proper pairs percent {.tabset}

### Uniquely mapped vs Proper Pairs Percent

```{r}
ggplot(as.data.frame(coldataBronch), aes(x = STAR_uniquely_mapped, y = RSeQC_proper_pairs_percent, color = Number_Outliers_Binned)) +
  geom_point(size = 0.75) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 14),
    axis.title = ggplot2::element_text(size = 14)
  ) +
  ggplot2::xlab(paste0("STAR Uniquely mapped")) +
  ggplot2::ylab(paste0("RSeQC Proper Pairs (%)")) +
  ggplot2::theme(legend.title = ggplot2::element_text(size = 15), legend.text = ggplot2::element_text(size = 13)) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 2)))
```

### Proper Pairs Percent vs TIN median

```{r}
ggplot(as.data.frame(coldataBronch), aes(x = RSeQC_proper_pairs_percent, y = RSeQC_TIN_median, color = Number_Outliers_Binned)) +
  geom_point(size = 0.75) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 14),
    axis.title = ggplot2::element_text(size = 14)
  ) +
  ggplot2::xlab(paste0("RSeQC Proper Pairs (%)")) +
  ggplot2::ylab(paste0("RSeQC TIN median")) +
  ggplot2::theme(legend.title = ggplot2::element_text(size = 15), legend.text = ggplot2::element_text(size = 13)) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 2)))

```

## PCA, QC outliers {.tabset}

We used `r isNumeric` QC metrics and computed the PCA on z-scored values. We subsequently ran Pearson correlation test on PC 1-10 and the z-score of each QC metric to calculate which principal component each QC metric was most strongly associated to.

### Outliers

```{r}
qcMetrics <- colDataBronchSubset

qcMetrics.z <- t(scale(qcMetrics))
qcMetrics.z.trim <- qcMetrics.z

qcMetrics.z.trim[qcMetrics.z < -2] <- -2
qcMetrics.z.trim[qcMetrics.z > 2] <- 2


RIN <- annotBronch$RIN
Number_Outliers_Binned <- as.factor(qualAnnot)
Number_Outliers_Binned <- factor(Number_Outliers_Binned, levels = c("x =< 25",
                                    "25 < x < 40",
                                    "x >= 40"))

Novartis_QC <- novartisQC

z <- qcMetrics.z

# run PCA
pc <- prcomp(z, center = F, scale = F)

forpcaBronch <- as.data.frame(pc$rotation[, 1:3])

forpcaBronch$Number_Outliers_Binned <- Number_Outliers_Binned

spc <- summary(pc)
pc1Importance <- signif(spc$importance[2, 1] * 100, 3)
pc2Importance <- signif(spc$importance[2, 2] * 100, 3)

# pdf("~/PCA_Batch.pdf", width = 9)
ggplot(forpcaBronch, aes(x = PC1, y = PC2, color = Number_Outliers_Binned)) +
  geom_point(size = 0.75) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 14),
    axis.title = ggplot2::element_text(size = 14)
  ) +
  ggplot2::xlab(paste0("PC1 (", pc1Importance, "%)")) +
  ggplot2::ylab(paste0("PC2 (", pc2Importance, "%)")) +
  ggplot2::theme(legend.title = ggplot2::element_text(size = 15), legend.text = ggplot2::element_text(size = 13)) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 2)))
```
```{r}
cor1 <- cor(pc$rotation[,1:10], t(qcMetrics.z), method = "pearson")

cor1Res <- apply(cor1, 2, function(x){
    return(names(which.max(abs(x))))
})

uniquePCs <- unique(cor1Res)[order(unique(cor1Res))]
qcpc.list = list()
for(x in 1:length(uniquePCs)){
  pc <- uniquePCs[x]
  qcpc.list[[pc]] = names(cor1Res)[cor1Res == pc]
}

QCPC.table <- plyr::ldply(qcpc.list, cbind)
colnames(QCPC.table) <- c("PC", "Metric")

QCPC.table %>%
  knitr::kable(
    format = "html", align = rep("c", ncol(QCPC.table) + 1),
    row.names = TRUE
  ) %>%
  kableExtra::kable_styling() %>%
  kableExtra::scroll_box(height = "250px")

```

Of all of the QC metrics, `r length(cor1Res[cor1Res == "PC1"])` were most associated, either positively or negatively, to PC1.

### Outliers, Novartis

```{r}
forpcaBronch$Quality <- Novartis_QC

# pdf("~/PCA_Batch.pdf", width = 9)
ggplot(forpcaBronch, aes(x = PC1, y = PC2, color = Quality)) +
  geom_point(size = 0.75) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 14),
    axis.title = ggplot2::element_text(size = 14)
  ) +
  ggplot2::xlab(paste0("PC1 (", pc1Importance, "%)")) +
  ggplot2::ylab(paste0("PC2 (", pc2Importance, "%)")) +
  ggplot2::theme(legend.title = ggplot2::element_text(size = 15), legend.text = ggplot2::element_text(size = 13)) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 2)))
```

## Heatmap, QC outliers

The numeric QC metric values were z-scored across samples, plotted by heatmap. A total of `r isNumeric` QC metrics were used.

```{r}

RIN <- annotBronch$RIN
Number_Outliers_Binned <- as.factor(qualAnnot)
Number_Outliers_Binned <- factor(Number_Outliers_Binned, levels = c("x =< 25",
                                    "25 < x < 40",
                                    "x >= 40"))

Novartis_QC <- novartisQC
annotation_col <- data.frame(RIN, Number_Outliers_Binned, Novartis_QC)

rownames(annotation_col) <- colnames(qcMetrics.z.trim)

annotation_row <- data.frame(Correlated_PC = cor1Res)
rownames(annotation_row) <- rownames(qcMetrics.z.trim)

callback <- function(hc, mat) {
  sv <- svd(t(mat))$v[, 1]
  dend <- reorder(as.dendrogram(hc), wts = sv)
  as.hclust(dend)
}

out <- pheatmap(qcMetrics.z.trim,
  color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(100),
  annotation_col = annotation_col,
  annotation_row = annotation_row,
  legend = T,
  show_rownames = F,
  show_colnames = F,
  fontsize_row = 9,
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  clustering_callback = callback
)

out

##Added 10/26/20. Used cutree to determine which samples to remove

clusterLabels <- sort(cutree(out$tree_col, k=2))
write.table(x = names(clusterLabels[clusterLabels == 2]),
        file = "20201026_PoorQuality_Bronch.txt", sep = "\t")

removed <- names(clusterLabels[clusterLabels == 2])
```

We split the dendrogram into 2 "branches" and considered all samples within the branch enriched for poor QC metrics to be of poor quality. These `r length(removed)` samples are the following: `r removed`


## PCA, Batch Effect {.tabset}

PCA was conducted on TPM counts. No evidence of batch effect can be seen in the PCA plot.

### PC1 vs PC2

```{r, warning = F, message = F}
counts <- assays(seBronch)[[2]]

counts <- counts[rowSums(counts) > 0, ]

# z normalize
z <- t(scale(t(counts)))
# run PCA
pc <- prcomp(z, center = F, scale = F)

forpcaBronch <- as.data.frame(pc$rotation[, 1:3])

forpcaBronch$Batch <- annotBronch$Batch

spc <- summary(pc)
pc1Importance <- signif(spc$importance[2, 1] * 100, 3)
pc2Importance <- signif(spc$importance[2, 2] * 100, 3)
pc3Importance <- signif(spc$importance[2, 3] * 100, 3)

# pdf("~/PCA_Batch.pdf", width = 9)
ggplot(forpcaBronch, aes(x = PC1, y = PC2, color = Batch)) +
  geom_point(size = 0.75) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 14),
    axis.title = ggplot2::element_text(size = 14)
  )  +
  ggplot2::xlab(paste0("PC1 (", pc1Importance, "%)")) +
  ggplot2::ylab(paste0("PC2 (", pc2Importance, "%)")) +
  ggplot2::theme(legend.title = ggplot2::element_text(size = 15), legend.text = ggplot2::element_text(size = 13)) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 2)))
# dev.off()
```

### PC1 vs PC3

```{r, warning = F, message = F}
# PCA: PC1 vs PC3
ggplot(forpcaBronch, aes(x = PC1, y = PC3, color = Batch)) +
  geom_point(size = 0.75) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 14),
    axis.title = ggplot2::element_text(size = 14)
  )  +
  ggplot2::xlab(paste0("PC1 (", pc1Importance, "%)")) +
  ggplot2::ylab(paste0("PC3 (", pc3Importance, "%)")) +
  ggplot2::theme(legend.title = ggplot2::element_text(size = 15), legend.text = ggplot2::element_text(size = 13)) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 2)))
```

### QC outliers

```{r, warning = F, message = F}
forpcaBronch <- cbind(forpcaBronch, qc.tally.gg)
ggplot(forpcaBronch, aes(x = PC1, y = PC2, color = Number_Outliers_Binned)) +
  geom_point(size = 0.75) +
  geom_point(
    data = subset(forpcaBronch, Number_Outliers_Binned == "x => 25"),
    aes(x = PC1, y = PC2, color = Number_Outliers_Binned)
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 14),
    axis.title = ggplot2::element_text(size = 14)
  )  +
  ggplot2::xlab(paste0("PC1 (", pc1Importance, "%)")) +
  ggplot2::ylab(paste0("PC2 (", pc2Importance, "%)")) +
  ggplot2::theme(legend.title = ggplot2::element_text(size = 15), legend.text = ggplot2::element_text(size = 13)) +
  ggplot2::guides(colour = ggplot2::guide_legend(override.aes = list(size = 2)))
```

## Sex status

```{r, warning = F, message = F}
sex.genes <- c("KDM5D", "UTY", "DDX3Y", "USP9Y", "TXLNG2P", "RPS4Y1", "XIST")
sex.genes.id <- match(sex.genes, rowRanges(seBronch)$external_gene_id)

Gender <- annotBronch$PERSON_GENDER
Sample_Type <- annotBronch$`Sample Type`
Gender <- gsub(" Gender", "", Gender)
annotation_col <- data.frame(Gender, Sample_Type)

counts <- assays(seBronch)$TPM
rownames(counts) <- rowRanges(seBronch)$external_gene_id

genes1 <- counts[sex.genes.id, ]

genes1.z <- t(scale(t(genes1)))
genes1.z.trim <- genes1.z
genes1.z.trim[genes1.z < -2] <- -2
genes1.z.trim[genes1.z > 2] <- 2

rownames(annotation_col) <- colnames(genes1.z.trim)

callback <- function(hc, mat) {
  sv <- svd(t(mat))$v[, 1]
  dend <- reorder(as.dendrogram(hc), wts = sv)
  as.hclust(dend)
}

out <- pheatmap(genes1.z.trim,
  color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(100),
  annotation_col = annotation_col,
  legend = T,
  show_rownames = T,
  show_colnames = F,
  fontsize_row = 9,
  cluster_rows = FALSE,
  cluster_cols = TRUE,
  clustering_callback = callback
)

exprs.male <- which(genes1.z.trim["KDM5D", ] > 1)
record.female <- which(annotBronch$PERSON_GENDER == "Female Gender")
sex.discrepancy <- annotBronch$`Kit Number`[intersect(exprs.male, record.female)]

out
```
The patient expressing male sex genes, but annotated female, is `r sex.discrepancy`

## Demographics Comparison

General annotation data of patients were compared between those of which samples were removed due to outliers, and those that were retained.

```{r, warning = FALSE, message = FALSE}
seBronchQC <- seBronch[, !colnames(seBronch) %in% highAmountOutlier]
annotBronchQC <- annotBronch[!annotBronch$`RNA Sample ID` %in% highAmountOutlier, ]

seBronchQCOut <- seBronch[, colnames(seBronch) %in% highAmountOutlier]
annotBronchQCOut <- annotBronch[annotBronch$`RNA Sample ID` %in% highAmountOutlier, ]

lab <- c(
  rep("Retained", length(annotBronchQC$GRPB_SMK_STS)),
  rep("Removed", length(annotBronchQCOut$GRPB_SMK_STS))
)
# AGE
mean.age <- mean(annotBronchQC$AGE)
sd.age <- sd(annotBronchQC$AGE)

age <- paste0(signif(mean.age, 3), " +/- ", signif(sd.age, 3))

mean.age.rm <- mean(annotBronchQCOut$AGE)
sd.age.rm <- sd(annotBronchQCOut$AGE)

age.rm <- paste0(signif(mean.age.rm, 3), " +/- ", signif(sd.age.rm, 3))

age.pvalue <- signif(wilcox.test(annotBronchQC$AGE, annotBronchQCOut$AGE)$p.value, 3)

demo.table <- c(age, age.rm, age.pvalue)

# Gender
gstatus <- c(annotBronchQC$PERSON_GENDER, annotBronchQCOut$PERSON_GENDER)

male.ct <- sum(annotBronchQC$PERSON_GENDER == "Male Gender")
male.pct <- signif(male.ct / length(annotBronchQC$PERSON_GENDER), 3) * 100
male <- paste0("Male: ", male.ct, "(", male.pct, "%)")

male.ct.rm <- sum(annotBronchQCOut$PERSON_GENDER == "Male Gender")
male.pct.rm <- signif(male.ct.rm / length(annotBronchQCOut$PERSON_GENDER), 3) * 100
male.rm <- paste0("Male: ", male.ct.rm, "(", male.pct.rm, "%)")

gender.pvalue <- signif(chisq.test(table(lab, gstatus))$p.value, 3)

demo.table <- rbind(demo.table, c(male, male.rm, gender.pvalue))

# FEVFVC
mean.fevfvc <- mean(annotBronchQC$FEV_FVC)
sd.fevfvc <- sd(annotBronchQC$FEV_FVC)

fevfvc <- paste0(signif(mean.fevfvc, 3), " +/- ", signif(sd.fevfvc, 3))

mean.fevfvc.rm <- mean(annotBronchQCOut$FEV_FVC)
sd.fevfvc.rm <- sd(annotBronchQCOut$FEV_FVC)

fevfvc.rm <- paste0(signif(mean.fevfvc.rm, 3), " +/- ", signif(sd.fevfvc.rm, 3))

fevfvc.pvalue <- signif(wilcox.test(annotBronchQC$FEV_FVC, annotBronchQCOut$FEV_FVC)$p.value, 3)

demo.table <- rbind(demo.table, c(fevfvc, fevfvc.rm, fevfvc.pvalue))

# Smoking Status
smk <- c(annotBronchQC$GRPB_SMK_STS, annotBronchQCOut$GRPB_SMK_STS)

annotBronchQC$GRPB_SMK_STS <- gsub(" ", "_", annotBronchQC$GRPB_SMK_STS)
annotBronchQCOut$GRPB_SMK_STS <- gsub(" ", "_", annotBronchQCOut$GRPB_SMK_STS)

smkstatus.na <- sum(is.na(annotBronchQC$GRPB_SMK_STS == "Current_Smoker"))

smkstatus.ct <- sum(annotBronchQC[!is.na(annotBronchQC$GRPB_SMK_STS), ]$GRPB_SMK_STS == "Current_Smoker")
smkstatus.pct <- signif(smkstatus.ct / nrow(annotBronchQC[!is.na(annotBronchQC$GRPB_SMK_STS), ]), 3) * 100
smkstatus <- paste0("Current: ", smkstatus.ct, "(", smkstatus.pct, "%)")

smkstatus.ct.rm <- sum(annotBronchQCOut$GRPB_SMK_STS == "Current_Smoker")
smkstatus.pct.rm <- signif(smkstatus.ct.rm / length(annotBronchQCOut$GRPB_SMK_STS), 3) * 100
smkstatus.rm <- paste0("Current: ", smkstatus.ct.rm, "(", smkstatus.pct.rm, "%)")

smkstatus.pvalue <- signif(chisq.test(table(lab, smk))$p.value, 3)

demo.table <- rbind(demo.table, c(smkstatus, smkstatus.rm, smkstatus.pvalue))

rownames(demo.table) <- c("Age", "Gender", "FEV1/FVC", "Smoking Status")
colnames(demo.table) <- c("Retained", "Removed", "p-value")

demo.table %>%
  knitr::kable(
    format = "html", align = rep("c", ncol(demo.table) + 1),
    row.names = TRUE
  ) %>%
  kableExtra::kable_styling()
```


## Differential Expression, Smoking status {.tabset}

DESeq2 was run (Model: ~ Smoking Status) on both expression data prior to and after filtering possible outliers.

### Pre-filter

```{r, warning = FALSE, message = FALSE, eval=T}
counts <- assay(seBronch)
rownames(counts) <- rowRanges(seBronch)$external_gene_id

counts <- apply(counts, 1:2, function(x) {
  return(as.integer(x))
})

no.smk.status.ix <- which(is.na(annotBronch$GRPB_SMK_STS))

counts <- counts[, -no.smk.status.ix]
annotBronch <- annotBronch[-no.smk.status.ix, ]
annotBronch$GRPB_SMK_STS <- gsub(" ", "_", annotBronch$GRPB_SMK_STS)

dds <- DESeqDataSetFromMatrix(
  countData = counts,
  colData = annotBronch,
  design = ~GRPB_SMK_STS
)
dds$GRPB_SMK_STS <- relevel(dds$GRPB_SMK_STS, ref = "Former_Smoker")

keep <- rowSums(counts(dds)) >= 10
dds <- dds[keep, ]
dds <- DESeq(dds)
res <- results(dds)

resOrdered <- res[order(res$pvalue), ]

resOrdered <- as.data.frame(resOrdered)
resOrdered$nLogPVal <- -log10(resOrdered$padj)
resSig <- resOrdered %>%
  filter(padj < 0.05) %>%
  filter(abs(log2FoldChange) > 0.5)
numSig <- nrow(resSig)
numUp <- nrow(resSig[resSig$log2FoldChange > 0.5, ])
numDown <- nrow(resSig[resSig$log2FoldChange < -0.5, ])

labels <- sapply(rownames(resOrdered), function(x) {
  if (x %in% c("CYP1A1", "CYP1B1")) {
    return("red")
  } else {
    return("black")
  }
})

labeltxt <- sapply(rownames(resOrdered), function(x) {
  if (x %in% c("CYP1A1")) {
    return("CYP1A1")
  } else if (x %in% c("CYP1B1")) {
    return("CYP1B1")
  } else {
    return("")
  }
})

resOrderedCYP <- resOrdered[c("CYP1A1", "CYP1B1"), ]

ggplot(resOrdered, aes(x = log2FoldChange, y = nLogPVal)) +
  geom_point(size = 0.75) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 14),
    axis.title = ggplot2::element_text(size = 14)
  ) +
  ggplot2::theme(legend.title = ggplot2::element_text(size = 15), legend.text = ggplot2::element_text(size = 13)) +
  ggplot2::guides(
    colour =
      ggplot2::guide_legend(override.aes = list(size = 2))
  ) + geom_point(data = resSig, aes(x = log2FoldChange, y = nLogPVal), color = "Red",
                 size = 0.75) +
  geom_point(data = resOrderedCYP, aes(x = log2FoldChange, y = nLogPVal), color = "blue") +
  ggrepel::geom_text_repel(aes(x = log2FoldChange, y = nLogPVal), label = labeltxt) +
  ggplot2::labs(y = "-log10(Adjusted p-value)") +
  ggplot2::ggtitle(label = "Smoking status, Former vs Current") +
  ggplot2::theme(plot.title = ggplot2::element_text(
    hjust = 0.5,
    size = 18
  ))
```

```{r}
# genenameTable <- as.data.frame(as.matrix(rowRanges(seBronch)$external_gene_id))
# 
# genenameTable$ensembleGeneId <- rownames(assay(seBronch))
# 
# resOrderedPreFilter <- resOrdered[match(genenameTable$V1, rownames(resOrdered)),]
# rownames(resOrderedPreFilter) <- genenameTable$ensembleGeneId
# 
# rownames(limmaRes) <- gsub(pattern = "_at", "", rownames(limmaRes))
# 
# limmaRes <- limmaRes[match(genenameTable$ensembleGeneId, rownames(limmaRes)),]
# 
# genelistIntersect <- intersect(rownames(limmaRes), rownames(resOrderedPreFilter))
# 
# limmaRes <- limmaRes[genelistIntersect,]
# resOrderedPreFilter <- resOrderedPreFilter[genelistIntersect,]
# 
# logfcTable <- as.data.frame(cbind(limmaRes$logFC, resOrderedPreFilter$log2FoldChange))
# colnames(logfcTable) <- c("Steiling", "DECAMP")
# summary(lm(DECAMP ~ Steiling, logfcTable))
```


### Post-filter

```{r, warning = FALSE, message = FALSE, eval=T}
counts <- assay(seBronchQC)
rownames(counts) <- rowRanges(seBronch)$external_gene_id

counts <- apply(counts, 1:2, function(x) {
  return(as.integer(x))
})

annotBronchQC$GRPB_SMK_STS <- gsub(" ", "_", annotBronchQC$GRPB_SMK_STS)

no.smk.status.ix <- which(is.na(annotBronchQC$GRPB_SMK_STS))

counts <- counts[, -no.smk.status.ix]
annotBronchQC <- annotBronchQC[-no.smk.status.ix, ]

dds <- DESeqDataSetFromMatrix(
  countData = counts,
  colData = annotBronchQC,
  design = ~GRPB_SMK_STS
)
dds$GRPB_SMK_STS <- relevel(dds$GRPB_SMK_STS, ref = "Former_Smoker")
keep <- rowSums(counts(dds)) >= 10
dds <- dds[keep, ]
dds <- DESeq(dds)
res <- results(dds)
# res

resOrdered <- res[order(res$pvalue), ]
# resOrdered
resOrdered <- as.data.frame(resOrdered)
resOrdered$nLogPVal <- -log10(resOrdered$padj)
resSig <- resOrdered %>%
  filter(padj < 0.05) %>%
  filter(abs(log2FoldChange) > 0.5)
numSigQC <- nrow(resSig)
numUpQC <- nrow(resSig[resSig$log2FoldChange > 0.5, ])
numDownQC <- nrow(resSig[resSig$log2FoldChange < -0.5, ])

labeltxt <- sapply(rownames(resOrdered), function(x) {
  if (x %in% c("CYP1A1")) {
    return("CYP1A1")
  } else if (x %in% c("CYP1B1")) {
    return("CYP1B1")
  } else {
    return("")
  }
})

resOrderedCYP <- resOrdered[c("CYP1A1", "CYP1B1"), ]

ggplot(resOrdered, aes(x = log2FoldChange, y = nLogPVal)) +
  geom_point(size = 0.75) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(size = 14),
    axis.title = ggplot2::element_text(size = 14)
  ) +
  ggplot2::theme(legend.title = ggplot2::element_text(size = 15), legend.text = ggplot2::element_text(size = 13)) +
  ggplot2::guides(
    colour =
      ggplot2::guide_legend(override.aes = list(size = 2))
  ) + geom_point(data = resSig, aes(x = log2FoldChange, y = nLogPVal), color = "Red",
                 size = 0.75) +
  geom_point(data = resOrderedCYP, aes(x = log2FoldChange, y = nLogPVal), color = "blue") +
  ggrepel::geom_text_repel(aes(x = log2FoldChange, y = nLogPVal), label = labeltxt) +
  ggplot2::labs(y = "-log10(Adjusted p-value)") +
  ggplot2::ggtitle(label = "Smoking status, Former vs Current") +
  ggplot2::theme(plot.title = ggplot2::element_text(
    hjust = 0.5,
    size = 18
  ))

```

## {.unlisted .unnumbered}

A total of `r numSig`(`r numUp` Upregulated, `r numDown` Downregulated) and `r numSigQC`(`r numUpQC` Upregulated, `r numDownQC` Downregulated) were differentially expressed pre- and post- filter, respectively(padj < 0.05, abs(log2FC) > 0.5).

## Reproducibility

<details><summary>For reproducibility:</summary>

```{r, warning = F, message = F}
## datetime

Sys.time()
```

```{r, warning = F, message = F}
## session info

sessionInfo()
```
</details>
